\pdfbookmark[1]{Acknowledgments}{acknowledgments}

\begin{flushright}
\textsl{When you are studying any matter, or considering any philosophy,
ask yourself only, what are the facts and what is the truth that the
facts bear out.}\\
\textsl{ Never let yourself be diverted either by what you wish to
believe, or by what you think would have beneficent social effects
if it were believed.}\\
\textsl{ But look only, and solely, at what are the facts.}
\par\end{flushright}

\begin{flushright}
--- Bertrand Russell 
\par\end{flushright}

\bigskip{}


\begingroup
\let\clearpage\relax
\let\cleardoublepage\relax 


\chapter*{Acknowledgements}

First of all, I would like to express my most respectful gratitude
to my thesis advisors, Daniel Borgis and Luc Belloni, who have developed
the main theories used in this thesis and and whose great knowledge
of liquid theory as well as the genius way of thinking and explaining
have given me a solid guide for doing this research. I would also
like to thank them for their tireless work in correcting this manuscript.

I'm grateful to Maximilien Levesque, the other main developer of MDFT
who joined the supervision of my thesis. Thanks for sharing some results
and data that have proven useful to my research.

I wish to acknowledge all the great minds willing to evaluate and
give ideas about my work: \textcolor{red}{(noms des jurés)}, thank
you for agreeing to be part of the jury of this thesis.

As a master student issue from pure chemistry speciality, a lack of
knowledge about Informatics brought to me a lot of difficulties. I
would like to express my sincere gratitude to my colleagues Pierre
Kestener, Matthieu Haefele, and Yacine Ould-Rouis for their huge aid
in Informatics and very useful advices during this thesis.

This thesis was produced at\textit{ Maison de la Simulation, CEA Saclay},
financially supported by the scholarship \textcolor{red}{IDEX-CEA}.
I acknowledge all the organizations and staff that gave me the chance
to have this three-year experience.

I am also grateful to Thomas Wiggins for help in correcting the huge
amount of grammar faults in this manuscript.

I'm deeply indebted to my tutors during my bachelor and master, respectively
Mr. Hongwei Tan and Mrs. Michelle Gupta, whose clear logic and warm
encouragement gave me all that I needed to be in love with theoretical
chemistry.

Finally I would like to thank my father, who made the right decision
to send me here in France and support me in every aspect.

\endgroup
